Computation and comparison of the stable Northeastern US marine boundary layer

### 2020-11-30 Paper/methodology.tex
Writing more sections of the paper

\subsection{Measured offshore conditions}
\fxnote{Lawrence write this part}

@@

From the measured distribution of atmospheric stabilities, turbulent
kinetic energies (TKE), and wind speeds from the Cape Wind platform, a
specific set of conditions were chosen as targets for this
computational study.  A summary of the conditions for all stability
classes is given table \ref{tab:CapeWindMeasurements}. In this work we
focus on the three stable atmospheric conditions at wind speeds of
5m/s, 10m/s, and 15m/s, while the neutral and unstable conditions
where studied in previous work \cite{cheung2020large}.

To maintain consistency with the measured data, the turbulence
intensity (TI) is calculated using the TKE as
\begin{equation}
  \textrm{TI} =
  \frac{\sqrt{\frac{2}{3}\times\textrm{TKE}}}{\overline{U}_{horiz}} =
  \frac{\sqrt{\frac{1}{3}\left( \overline{u'u' + v'v' + w'w'}
      \right)}}{\overline{U}_{horiz}}
\end{equation}
The averaged wind speeds were enforced at the measurement height of
20m, and the wind directions...

@@

\subsection{Computational methodlogy}
Provide a general description of the LES codes that we use.

@@

\subsubsection{Lower wall model BC}
\fxnote{Ganesh write this part}\\
What we implemented in terms of the lower boundary condition.

@@

\subsection{Computational set up}
\fxnote{Lawrence write this part}\\
Domains, grids, boundary conditions

\begin{table}
\caption{\label{tab:Computational parameters} LES parameters for stable conditions}
\centering
\begin{tabular}{cccc}
  \hline
  Stability    & Wind speed  & Surface roughness $z_0$ & Temperature gradient \\
  \hline
  Stable       & 5  m/s           & 0.0005 m       & -0.32 K/hr      \\
  Stable       & 10 m/s           & 0.0005 m       & -1.40 K/hr      \\
  Stable       & 15 m/s           & 0.0005 m       & -1.50 K/hr      \\
\hline
\end{tabular}
\end{table}

\subsubsection{Simulation workflow}
\fxnote{Lawrence write this part}\\
Describe the process to hit target conditions.

>>>>

\subsection{Measured offshore conditions}
% \fxnote{Lawrence write this part}

@@

From the measured distribution of atmospheric stabilities, turbulent
kinetic energies (TKE), and wind speeds from the Cape Wind platform at
z=20m, a stable set of conditions were chosen as targets for this
computational study.  A summary of the conditions for all stability
classes is given table \ref{tab:CapeWindMeasurements}. In this work we
focus on the three stable atmospheric conditions at wind speeds of
5m/s, 10m/s, and 15m/s, while the neutral and unstable conditions
where studied in previous work \cite{cheung2020large}.

To maintain consistency with the measured data, the turbulence
intensity (TI) is calculated using the TKE as
\begin{equation}
  \textrm{TI} =
  \frac{\sqrt{\frac{2}{3}\times\textrm{TKE}}}{\overline{U}_{horiz}} =
  \frac{\sqrt{\frac{1}{3}\left( \overline{u'u' + v'v' + w'w'}
      \right)}}{\overline{U}_{horiz}}
\end{equation}
The averaged wind speeds were enforced at the measurement height of
20m, and the applied wind directions were consistent with the
predominant stable direction of 225 degrees southwest.

@@

\subsection{Computational methodology}
Provide a general description of the LES codes that we use.

@@

\subsubsection{\label{sec:wallmodelBC}Lower wall model BC}
\fxnote{Ganesh write this part}\\
What we implemented in terms of the lower boundary condition.

@@

\subsection{Computational setup}
% \fxnote{Domains, grids, boundary conditions}

\begin{table}
\caption{\label{tab:z0tempparam} LES parameters for stable ABL conditions}
\centering
\begin{tabular}{ccccc}
  \hline
  Stability & Wind speed & Surface roughness $z_0$ & Surface
  temperature change & timestep\\
  \hline
  Stable       & 5  m/s           & 0.0005 m       & -0.32 K/hr   & 0.25 sec   \\
  Stable       & 10 m/s           & 0.0005 m       & -1.40 K/hr   & 0.125 sec  \\
  Stable       & 15 m/s           & 0.0005 m       & -1.50 K/hr   & 0.0625 sec \\
\hline
\end{tabular}
\end{table}

The computational methodology for the AMR-Wind and Nalu-Wind LES codes
follows practices similar to the previous offshore ABL study, with
some modifications to handle the stable stratification.  As in
\cite{cheung2020large}, the domain was a square prism geometry, with
the $x$ and $y$ coordinates aligned in the East and North directions,
respectively.  The domain size and mesh requirements were determined
through a numerical grid study.  As discussed in section
\ref{sec:gridstudy}, the horizontal dimensions of 750m $\times$ 750m
were found to be sufficient to capture any large scale structures in
the boundary layer, and in all cases the vertical dimension of 1000m
was also used.  The grid study in section \ref{sec:gridstudy} also
showed that the stable stratification required greater mesh refinement
to capture the smaller turbulent scales, and a uniform cell size of
2.5m was adopted for the offshore LES computations across both codes.

Momentum source terms were included in Nalu-Wind and AMR-Wind to
ensure that the horizontally averaged velocity matched the targeted
wind speed at the z=20 m height.  These source terms are based on the
difference between the desired wind velocity and the instantaneous
horizontally averaged velocity, and are only a function of time and
height z.  Coriolis forcing matching the Cape Wind latitude was
included to capture the effect of wind change with elevation.

\subsubsection{Boundary and initial conditions}
In both horizontal directions of the computational domain, periodic
boundary conditions were applied.  At the lower boundary, we chose to
represent the air/ocean interface using flat boundary with small
amount of surface roughness.  This allowed the wall model discussed in
section \ref{sec:wallmodelBC} to be applied.  Monin-Obukhov similarity
theory was used to determine the velocity and temperature profiles
near the lower surface given a surface roughness height $z_0$ and the
prescribed surface temperature as a function of time.  At the upper
surface of the domain, a potential flow based boundary condition is
applied along with a normal temperature gradient of 0.003 K/m.

The initial temperature profile in all cases was a constant 300K until
the specified inversion height of 650m.  The inversion layer thickness
was 100m, and above this, the temperature linearly increased until it
reached 308.75K at the upper boundary.  The initial mean velocity
profile was uniform throughout the domain with superimposed sinusoidal
velocity perturbations of magnitude 1 m/s to promote the development
of turbulence.

\subsubsection{Determination of surface roughness and prescribed temperature }
To match the measured TI conditions given in table
\ref{tab:CapeWindMeasurements}, the surface roughness and prescribed
surface temperature change were adjusted through an initial
trial-and-error process.  The final surface roughness of $z_0$=0.0005m
used the stable offshore conditions (see table \ref{tab:z0tempparam})
matched the roughness used in the neutral 15m/s and unstable 5m/s and
10m/s cases from the previous study \cite{cheung2020large}.  This
value of surface roughness is consistent with the measurements from
the North Sea \cite{taylor2001dependence}, which found values of $z_0$
ranging from $5 \times 10^{-5}$m to $5\times 10^{-3}$m.

The values of the prescribed temperature change at the ocean surface
is also given in table \ref{tab:CapeWindMeasurements}.  As the wind
speed increases, a larger temperature decrease was used to maintain
similar levels of stable stratification, as expected from
Monin-Obukhov similarity theory.

### 2020-11-30 Paper/results.tex
Writing more sections of the paper

\subsection{Domain and grid resolution study}

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\caption{\label{tab:GridStudySetup} The setup for the grid resolution study} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & dx [m] & dy [m] & dz [m] \\
  \hline
  Nalu-wind coarse  &  10.0  & 10.0   & 2.5   \\
  Nalu-wind medium  &   5.0  &  5.0   & 2.5   \\
  Nalu-wind fine    &   2.5  &  2.5   & 2.5   \\
  AMR-wind fine     &   2.5  &  2.5   & 2.5   \\
\hline
\end{tabular}
\end{table}

@@

@@

%%%%%%%%%%% Grid resolution spectra figure %%%%%%%%%%%%%%%%%%%%%%%%%
% Created in Postprocessing/GridStudy/GridStudy_Spectra.ipynb
\begin{figure}[hbt!]
  \label{fig:GridStudySpectra}
  \centering
  \includegraphics[width=2.0in]{figures/GridStudy_Spectra_Su.png}
  \includegraphics[width=2.0in]{figures/GridStudy_Spectra_Sv.png}
  \includegraphics[width=2.0in]{figures/GridStudy_Spectra_Sw.png}
  \caption{Calculation of the wind spectra $S_i$ for LES of stable
    5m/s case with different resolutions.  The gray vertical lines
    correspond to the maximum resolvable frequency according to
    equation (\ref{eq:fmax}). }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%% Grid resolution correlation figure %%%%%%%%%%%%%%%%%%%%
% Created in Postprocessing/GridStudy/GridStudy_Rij.ipynb
\begin{figure}[hbt!]
  \label{fig:GridStudyRij}
  \centering
  \includegraphics[width=3in]{figures/GridStudy_Rij_Longitudinal.png}
  \includegraphics[width=3in]{figures/GridStudy_Rij_Lateral.png}
  \caption{Calculation of the averaged longitudinal and lateral
    $R_{ij}(\xi)$ coefficient at $z$=20m for LES of stable 5m/s case
    with different resolutions.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\caption{\label{tab:GridStudyLscale} The calculated turbulent integral
  lengthscale for each of the grid resolutions} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & dx [m] & Longitudinal L [m] & Lateral L [m] \\
  \hline
  Nalu-wind coarse  &  10.0  & 36.433999          & 0.0        \\
  Nalu-wind medium  &   5.0  & 21.485711          & 4.519247   \\
  AMR-wind fine     &   2.5  & 27.689340          & 9.332672   \\
\hline
\end{tabular}
\end{table}

\subsection{Comparison of AMR-Wind vs Nalu-Wind}

@@

The Kaimal model for spectra \cite{kaimal1973turbulence, cheynet2017spectral}
\begin{equation}
  \label{eq:kaimal}
  \frac{fS_i}{u_\tau^2} = \frac{a_i(fz/\bar{U})}{\left(1+b_i(fz/\bar{U})^{\alpha_i}\right)^{\beta_i}}
\end{equation}

@@

@@

>>>>

\subsection{\label{sec:gridstudy}Domain and grid resolution study}

In order to ensure that the large and small turbulence scales were
properly resolved in the offshore LES simulations, the appropriate
domain size and mesh resolution needed to be determined.
Under-resolved meshes or domain sizes which are too small may miss the
smaller turbulence scales or remove the larger structures required for
proper ABL development.  However, over-resolved meshes or
unnecessarily large domains might lead to excessive computational
requirements to complete the simulations.

The appopriateness of these meshes and domain sizes were determined
based on wind spectra and integral length scale they were able to capture.

@@

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\caption{\label{tab:GridStudySetup} The setup for the grid resolution study} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & dx [m] & dy [m] & dz [m] & Domain size \\
  \hline
  Nalu-wind coarse  &  10.0  & 10.0   & 2.5 & 1.5km x 1.5km x 1.0km  \\
  Nalu-wind medium  &   5.0  &  5.0   & 2.5 & 1.5km x 1.5km x 1.0km  \\
  Nalu-wind fine    &   2.5  &  2.5   & 2.5 & 0.0 \\
  AMR-wind fine     &   2.5  &  2.5   & 2.5 & 0.0 \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%% Grid resolution spectra figure %%%%%%%%%%%%%%%%%%%%%%%%%
% Created in Postprocessing/GridStudy/GridStudy_Spectra.ipynb
\begin{figure}[hbt!]
  \label{fig:GridStudySpectra}
  \centering
  \includegraphics[width=2.0in]{figures/GridStudy_Spectra_Su.png}
  \includegraphics[width=2.0in]{figures/GridStudy_Spectra_Sv.png}
  \includegraphics[width=2.0in]{figures/GridStudy_Spectra_Sw.png}
  \caption{Calculation of the wind spectra $S_i$ for LES of stable
    5m/s case with different resolutions.  The gray vertical lines
    correspond to the maximum resolvable frequency $f_{max}$ according
    to equation (\ref{eq:fmax}). }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%% Grid resolution Rij correlation figure %%%%%%%%%%%%%%%%%%%%
% Created in Postprocessing/GridStudy/GridStudy_Rij.ipynb
\begin{figure}[hbt!]
  \label{fig:GridStudyRij}
  \centering
  \includegraphics[width=3in]{figures/GridStudy_Rij_Longitudinal.png}
  \includegraphics[width=3in]{figures/GridStudy_Rij_Lateral.png}
  \caption{Calculation of the averaged longitudinal and lateral
    $R_{ij}(\xi)$ coefficient at $z$=20m for LES of stable 5m/s case
    with different resolutions.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\caption{\label{tab:GridStudyLscale} The calculated turbulent integral
  lengthscale for each of the grid resolutions} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & dx [m] & Longitudinal L [m] & Lateral L [m] \\
  \hline
  Nalu-wind coarse  &  10.0  & 36.433999          & 0.0        \\
  Nalu-wind medium  &   5.0  & 21.485711          & 4.519247   \\
  Nalu-wind fine    &   2.5  & 25.250568          & 5.919645   \\
  AMR-wind fine     &   2.5  & 27.689340          & 9.332672   \\
\hline
\end{tabular}
\end{table}

\subsection{Comparison of AMR-Wind vs Nalu-Wind}
\fxnote{COMPARE AMR-WIND VS NALU-WIND HERE}

%%%%%%%%%%%%%%% Compare Nalu/AMR integrated quantities %%%%%%%%%%%%%%
\begin{table}
\caption{\label{tab:CompareAMRvsNalu} Comparison of AMR-Wind and Nalu-Wind} \centering
\begin{tabular}{ccccc}
  \hline
  Case          & TI        &  $\alpha$  &   $u_\tau$ &       L \\
  \hline
  naluwind 5m/s &  0.048106 &  0.165277 &  0.163342 & 94.970836 \\
  amrwind 5m/s  &  0.048255 &  0.165552 &  0.157399 & 52.512434 \\
  \hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%% Compare Nalu/AMR profiles %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AMRWind_NaluWind_stable_05ms_mesh2_5x2_5x2_5.ipynb
\begin{figure}[hbt!]
  \label{fig:CompareAMRvsNaluProfiles}
  \centering
  \includegraphics[width=6.0in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_stable_05ms_mesh2p5_2p5_2p5_WS.png}\\
  \includegraphics[width=6.0in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_stable_05ms_mesh2p5_2p5_2p5_Wdir.png}\\
  \includegraphics[width=6.0in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_stable_05ms_mesh2p5_2p5_2p5_TI.png}
  \caption{Comparison of AMR-Wind and Nalu-Wind profiles. }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%% Compare Nalu/AMR spectra %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Postprocessing/ABLSpectra/AMRWind_NaluWind_Spectra_Stable.ipynb
\begin{figure}[hbt!]
  \label{fig:CompareAMRvsNaluSpectra}
  \centering
  \includegraphics[width=6.0in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_Spectra_Stable_z20.png}

  \caption{Comparison of AMR-Wind and Nalu-Wind wind spectra at
    z=20m. }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%% Compare Nalu/AMR lengthscale %%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Postprocessing/ABLLength/CompareAMRNalu_ABL_Lengthscales.ipynb
\begin{figure}[hbt!]
  \label{fig:CompareAMRvsNaluRij}
  \centering
  \fxnote{UPDATE THIS FIGURE}\\
  \includegraphics[width=2.5in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_Lengthscale_Stable_z20_Longitudinal.png}
  \includegraphics[width=2.5in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_Lengthscale_Stable_z20_Lateral.png}

  \caption{Comparison of AMR-Wind and Nalu-Wind $\langle R_{ij}
    \rangle$ correlation at z=20m. }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The Kaimal model for spectra \cite{kaimal1973turbulence,
  cheynet2017spectral} for neutral atmospheric conditions:
\begin{equation}
  \label{eq:kaimal}
  \frac{fS_i}{u_\tau^2} = \frac{a_i(fz/\bar{U})}{\left(1+b_i(fz/\bar{U})^{\alpha_i}\right)^{\beta_i}}
\end{equation}

@@

%%%%%%%%%%% Grid resolution spectra figure %%%%%%%%%%%%%%%%%%%%%%%%%
% Created in Postprocessing/ABLSpectra/Stable_Spectra_Allz.ipynb
\begin{figure}[hbt!]
  \label{fig:ABLSpectra_AllZ}
  \centering
  \includegraphics[width=7.0in]{figures/Stable_Spectra_AllZ_05ms.png}\\
  \includegraphics[width=7.0in]{figures/Stable_Spectra_AllZ_10ms.png}\\
  \includegraphics[width=7.0in]{figures/Stable_Spectra_AllZ_15ms.png}
  \caption{Calculation of the wind spectra $S_i$ for LES of stable
    5m/s, 10m/s, and 15m/s cases at z=20m, 40m, and 60m.  The black
    vertical lines correspond to the maximum resolvable frequency
    $f_{max}$ according to equation (\ref{eq:fmax}). }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\caption{\label{tab:StabilityStudyLscale} The calculated turbulent
  integral lengthscale for each of atmospheric stabilities} \centering
\begin{tabular}{ccccc}
  \hline
  Stability   & Wind speed & Longitudinal L [m] & Lateral L [m] \\
  \hline
  Stable      &   5 m/s  & 0.0           & 0.0        \\
  Stable      &  10 m/s  & 0.0           & 0.0        \\
  Stable      &  15 m/s  & 0.0           & 0.0        \\
  Neutral     &   5 m/s  & 110.435741    & 15.245327  \\
  Neutral     &  10 m/s  & 165.398368    & 20.634052  \\
  Neutral     &  15 m/s  & 184.081967    & 22.961169  \\
  Unstable    &   5 m/s  & 187.868710    & 270.538753 \\
  Unstable    &  10 m/s  & 177.457502    & 215.027457 \\
  Unstable    &  15 m/s  & 263.475309    & 67.999898  \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
